---
output: html_document
---
# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
setwd("~/GitHub/RepData_PeerAssessment1")
activity <- read.csv("activity.csv", colClasses = c("integer","Date","integer"))
```

## What is mean total number of steps taken per day?
For this part of the assignment, you can ignore the missing values in
the dataset.

**1.** Make a histogram of the total number of steps taken each day

```{r}
library(dplyr)
library(ggplot2)
qplot(date, steps, data = activity %>% group_by(date), geom = "bar", stat = "identity", main = "Total number of steps per day (NAs removed)")
```

**2.** Calculate and report the **mean** and **median** total number of steps taken per day

```{r}
tot_steps <- activity %>% group_by(date) %>% summarise(tot_steps = sum(steps, na.rm=TRUE))
avg_steps <- mean(tot_steps$tot_steps, na.rm=TRUE)
med_steps <- median(tot_steps$tot_steps, na.rm=TRUE)
```
Mean number of steps per day: `r avg_steps`
Median number of steps perday: `r med_steps`

## What is the average daily activity pattern?

**1.** Make a time series plot (i.e. `type = "l"`) of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r Plot average daily activity}
qplot(interval, steps, data = activity %>% group_by(interval) %>% summarise(steps= mean(steps, na.rm=TRUE)), geom = "line", main = "Averge number of steps for 5 minute intervals")
```

**2.** Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
d1 <- activity %>% group_by(interval) %>% summarise(avg_steps = mean(steps, na.rm=TRUE))
with(d1,d1[order(-avg_steps),])
```
The interval with the maximum average number of steps is **835**

## Imputing missing values

Note that there are a number of days/intervals where there are missing
values (coded as `NA`). The presence of missing days may introduce
bias into some calculations or summaries of the data.

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with `NA`s)
```{r}
summary(activity)
```
There are **2304 `NA`s**

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

#### Replace missing values with the daily mean
```{r}
act_nareplace <- activity #Duplicate original data frame
#Replace values with mean number of steps over all observations
act_nareplace$steps[is.na(act_nareplace$steps)] <- mean(act_nareplace$steps, na.rm = TRUE)
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the **mean** and **median** total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}
qplot(date, steps, data = act_nareplace %>% group_by(date), geom = "bar", stat = "identity", main = "Total number of steps per day (NAs replaced)")

summary(act_nareplace %>% group_by(date) %>% summarise(tot_steps = sum(steps)))
```
New mean = 10766
New median = 10766

## Are there differences in activity patterns between weekdays and weekends?

```{r}
act_nareplace$weekday <- weekdays(act_nareplace$date)
act_nareplace$weekend <- act_nareplace$weekday=="Saturday" | act_nareplace$weekday=="Sunday"
act_nareplace$weekend <- ifelse(act_nareplace$weekend==TRUE,"Weekend","Weekday")

qplot(interval, steps, data = act_nareplace %>% group_by(interval,weekend) %>% summarise(steps = mean(steps)), facets = weekend~., geom = "line", ylab = "Number of steps")
```